# Introduction to R

## Arithmetic operations

- Using R as calculator
- Examples with [arithmetic operators](https://www.tutorialgateway.org/r-arithmetic-operators/)

```{r}
# Example for R Arithmetic Operators
a <- 16
b <- 3
add <- a + b
sub = a - b
multi = a * b
division = a / b
Integer_Division = a %/% b
exponent = a ^ b
modulus = a %% b

print(paste("Addition of two numbers 16 and 3 is : ", add))
print(paste("Subtracting Number 3 from 16 is : ", sub))
print(paste("Multiplication of two numbers 16 and 3 is : ", multi))
print(paste("Division of two numbers 16 and 3 is : ", division))
print(paste("Integer Division of two numbers 16 and 3 is : ", Integer_Division))
print(paste("Exponent of two numbers 16 and 3 is : ", exponent))
print(paste("Modulus of two numbers 16 and 3 is : ", modulus))
```

The book text demonstrates only some of the arithmetic operators.

Everything is an object/function! For instance:

```{r ops}
`+`(2,5)
```

Arithmetic operators are just one group of five different types of operators:

- Arithmetic operators
- Assignment operators
- Comparison operators
- Logical operators
- Miscellaneous operators

See compilation of operators at [tutorialspoint](https://www.tutorialspoint.com/r/r_operators.htm), [datamentor](https://www.datamentor.io/r-programming/operator/) and [w3schools](https://www.w3schools.com/r/r_operators.asp).

And what about the pipe operator?

::: {.infobox}
The challenge: How to explain some (most) of the operators during an ongoing research investigation? 
The remainder of unexplained operators should be addressed with the interactive tutorials.
:::

- Practical usage of the assignment operator(s) is easy because we need it all the time.
- Comparison and logical operators are to be explained with the `filter()` command. 
- But what could be a practical example with arithmetic operators? A calculation inside mutate to generate a new variable/column? I do not need all of them. Maybe +, -, *, / would be for the first take enough? But it shouldn't be forced. Perhaps the usage as demonstration for R as calculator is enough? It would be a good start to become acquainted with R.

## Objects

- Naming objects
- Assigning values to objects —> assignment operator(s)
- `class()` function
- Printing the value of objects 
    - just the name,
        - naming rules
        - case sensitivity
        - camelCase versus snake_case
        - see [slides by Jenny Bryan](https://www2.stat.duke.edu/~rcs46/lectures_2015/01-markdown-git/slides/naming-slides/naming-slides.pdf)
    - parentheses around the function
    - `print()` function
    - `cat()` 
        - outputing objects, 
        - concatenating the representations
        - performs much less conversion than `print()`.
    - `glue()` from the {glue} package
        - expressions enclosed by braces will be evaluated as R code 
        - long strings are broken by line and concatenated together
        - leading whitespace and blank lines from the first and last lines are automatically trimmed

## Vectors

- demonstrate usefulness
- one-dimensional array simply represents a collection of information stored in a specific order
- function `c()`, c stands for for concatenate
- indexing of vectors
- working with vectors: add/multiply one value, another vector same length, sequencing with colon
etc.

    

## Functions

- using some basic functions: length, min, max, range, mean, sd, sum etc.
- using some other important functions: seq, names, 
- how to call functions: parameters, default values etc.
- create functions: example of special summary-function.

## Data files

- `read_csv()` instead of `read.csv()` (other of this family)
- reading data in with RStudio -> get code (for reproducibility!)
- `here()` function instead of setting working directory with `setwd()`
- RStudio projects
- understanding data
    - str()
    - using tibble() instead of data.frame()
    - names(), nrow(), ncol(), dim()
    - glimpse() 
    - random selection of some rows by using `slice_sample()`
    - skim() of {skimr} as alternative to summary()


## Saving objects

- `save.image()` -> **NO!**
- save(), write_delim family e.g., `write_csv()`
- configure RStudio to prevent saving image
- `saveRDS()`, `readRDS()` instead of `save()` and `load()`: See J[enny Bryan's stat545](https://stat545.com/import-export.html)

## Packages

- `install.packages("file_name")`
- `library(file_name)` 

## Prgramming & learning tips

- script: `source()`
- RMarkdown: YAML header, chunks, chunks labels
- workflow: different kinds of executing program code
- shortcuts for important workflow actions
- chunk names, incl. reserved "setup" name

